Data-Driven Modelling \& \\ Forecasting of Chaotic Systems

### 2022-01-18 projectReport/chapter3.tex
checkpoint

\section{Practical Limitations to Takens' Theorem}
Takens' Embedding Theorem is a powerful result, and it provides compelling reason to believe that one could conceivably reconstruct a system conjugate accurately to the underlying system. 
Nonetheless, it does present some severe practical limitations:
\vspace{-5mm}
\begin{enumerate}
\item Even supposing that we could indeed find $F_\theta$, our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor. There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
\item Takens theorem is stated only for noiseless observations. Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e. the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely. This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter\ref{ch3}. 
\end{enumerate}

>>>>

\section{Practical Limitations to Takens' Theorem}\label{sect_takenslimits}
Takens' Embedding Theorem is a powerful result, and it provides compelling reason to believe that one could conceivably reconstruct a system conjugate accurately to the underlying system. 
Nonetheless, it does present some severe practical limitations:
\vspace{-5mm}
\begin{enumerate}
\item Even supposing that we could indeed find $F_\theta$, our \emph{approximation} of  $F_\theta$ is a map from a larger set $\mathbb{R}^{2d+1}$ containing the embedded attractor. There are, however, no theoretical guarantees that $F_\theta$ will retain  $V=\Phi_{2d,\theta}(W)$  as an attractor although $W$ could itself be an attractor.
\item Takens theorem is stated only for noiseless observations. Due to noise $\epsilon_n$ the delay vector $\Phi_{2d,\theta}(w_n) + \epsilon_n$  may lie outside $V$.Furthermore, due to the chaotic nature of the underlying system (i.e. the fact that it has SIDC), the evolution of $\Phi_{2d,\theta}(w_n) + \epsilon_n$ under the map $F_\theta$ could move out of $V$ completely. This problem can be overcome  by using a driven dynamical system with some properties, and we discuss this in Chapter\ref{ch3}. 
\end{enumerate}

### 2022-01-18 projectReport/chapter5.tex
checkpoint

Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not for a lossy approximation since we use all principal components \textbf{(Cite.)}\ednote{Ask M: Is there a specific resource to cite?}.
More explicitly, denote $X_{1:N}$ as the matrix with the first $N$ states of the network data represented by row vectors. If $X_{1:N}=U\Sigma P^T$\ednote{Expand here} represents the singular value decomposition of $X_{1:N}$ \ednote{Ask: Is there a specific resource to cite?}, we then denote the principal component matrix $P$, and the principal components are given by 
 $Z_{1:N}=X_{1:N}P$.
These principal components are then used to train a feedforward neural network. Denote the row vectors of $Z_{1:N}$ by 
$z_i^T, i=1,2,\ldots,N$ and the neural network by $NN$, the network is subsequently trained to learn an approximation of the map
$NN: (z_{n-1},z_n) \mapsto  u_n.
$
The learnt approximation of $NN$ can be used to approximate $\Gamma$ since 
\[
NN\left( \begin{bmatrix} 
P^Tx_{n-1} \\
P^Tx_n
\end{bmatrix}
\right) = NN \circ 
\begin{bmatrix}
P^T & 0 \\
0 & P^T 
\end{bmatrix}\begin{bmatrix}
x_{n-1}\\
x_n
\end{bmatrix} = u_n = \Gamma(x_{n-1},x_n).
\]

@@

\begin{center}
  \begin{table} 
      \scalebox{0.75}
      {\begin{tabular}{|c|c| c c |} 
          \toprule
          Input & Dimension of $X$ & $u_n$ vs $u_{n+1}$ 
          & $\begin{bmatrix}
              x_{n-1}\\
              x_n
          \end{bmatrix}$ vs $\begin{bmatrix}
              x_n\\
              x_{n+1}
          \end{bmatrix}$ \\
          \midrule      
          % \hline
          % $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
          % \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
          %     & 100 &  & 0.9934\\
          %     & 1000 & & 0.9930\\              
          % \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
          %     & 10 &0.8401  & 0.9392\\
          %     & 100 & & 0.9616\\
          %     & 1000 & & 0.9737\\
          % \hline

               \midrule  
               $(w_n)$ comes from the Clifford Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
        {$w_{n+1}= \begin{bmatrix} \sin(aw_{n,y}) + c\cos(aw_{n,x}) \\ 
                                                                    \sin(bw_{n,x})+d\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
              & 10 & -0.21713 & 0.78719 \\
              & 100 & &  0.8334 \\
              & 1000 & & 0.84854 \\
          % \hline 
          % $(w_n)$ comes from the Logistic map: & & & \\
          % {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
          % \multirow{3}{*}{$u_n = w_n - 0.5$} 
          %     & 10 &-0.0314  & 0.5577\\
          %     & 100 &  & 0.7150\\
          %     & 1000 & & 0.7826\\
          % \hline
          % \hline
          % $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
          % {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
          %                                                             c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
          %   \multirow{3}{*}{$u_n = w_n-\overline{w}$}
          %       & 10 & -0.2243 & 0.49698 \\
          %       & 100 & &  0.63310 \\
          %       & 1000 & & 0.71151 \\
           \bottomrule
                \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
 \caption{Following~\cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN. Rows correspond to distinct dynamical systems considered in the experiments
      The second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
      The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship $\begin{bmatrix}
              x_{n-1}\\
              x_n
          \end{bmatrix}$ vs $G_T \left( \begin{bmatrix}
              x_{n-1}\\
              x_n
          \end{bmatrix}\right)$ may be contrasted by the reader. 
          Finally, the RNN's dimension (from~\eqref{Seq_RNN}) is provided to exhibiting that an increase in dimension of the driven system will 
          typically result in a map $G_T$ with less functional complexity.}.
      \end{table}\label{tbl_attractorsPearson}
\end{center}

>>>>

Principal components are used only for a more efficient state representation to possibly reduce learning errors; it is not for a lossy approximation since we use all principal components \textbf{(Cite.)}\ednote{Ask M: Is there a specific resource to cite?}.
More explicitly, denote $X_{1:N}$ as the matrix with the first $N$ states of the network data represented by row vectors. If $X_{1:N}=U\Sigma P^T$\ednote{Expand here} represents the singular value decomposition of $X_{1:N}$ \ednote{Ask: Is there a specific resource to cite?}, we then denote the principal component matrix $P$, and the principal components are given by 
 $Z_{1:N}=X_{1:N}P$.
These principal components are then used to train a feedforward neural network. Denote the row vectors of $Z_{1:N}$ by 
$z_i^T, i=1,2,\ldots,N$ and the neural network by $NN$, the network is subsequently trained to learn an approximation of the map
$NN: (z_{n-1},z_n) \mapsto  u_n.
$
The learnt approximation of $NN$ can be used to approximate $\Gamma$ since 
\begin{equation}\label{Seq_RNN}
  NN\left( \begin{bmatrix} 
    P^Tx_{n-1} \\
    P^Tx_n
    \end{bmatrix}
    \right) = NN \circ 
    \begin{bmatrix}
    P^T & 0 \\
    0 & P^T 
    \end{bmatrix}\begin{bmatrix}
    x_{n-1}\\
    x_n
    \end{bmatrix} = u_n = \Gamma(x_{n-1},x_n).
\end{equation}

@@

\begin{center}
  \begin{table} 
      \scalebox{0.75}
      \centering
      {\begin{tabular}{|c|c| c c |} 
          \toprule
          Input & Dimension of $X$ & $u_n$ vs $u_{n+1}$ 
          & $\begin{bmatrix}
              x_{n-1}\\
              x_n
          \end{bmatrix}$ vs $\begin{bmatrix}
              x_n\\
              x_{n+1}
          \end{bmatrix}$ \\
          \midrule      
          % \hline
          % $(w_n)$ is the Lorenz states, sampled every 0.1 timestamps. & & &\\
          % \multirow{3}{*}{$u_n=\frac{1}{100}w_n$} & 10 &0.9311  & 0.9731\\
          %     & 100 &  & 0.9934\\
          %     & 1000 & & 0.9930\\              
          % \multirow{3}{*}{$u_n=\frac{1}{10}(\sin(0.1w_{n,x})+\sin(0.1w_{n,y})+\sin(0.1w_{n,z}))$} 
          %     & 10 &0.8401  & 0.9392\\
          %     & 100 & & 0.9616\\
          %     & 1000 & & 0.9737\\
          % \hline

               \midrule  
               $(w_n)$ comes from the Clifford Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
        {$w_{n+1}= \begin{bmatrix} \sin(aw_{n,y}) + c\cos(aw_{n,x}) \\ 
                                                                    \sin(bw_{n,x})+d\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
          \multirow{3}{*}{$u_n = w_n-\overline{w}$}
              & 10 & -0.21713 & 0.78719 \\
              & 100 & &  0.8334 \\
              & 1000 & & 0.84854 \\
          % \hline 
          % $(w_n)$ comes from the Logistic map: & & & \\
          % {$ w_{n+1}  = 4w_n(1-w_n)$} & & & \\
          % \multirow{3}{*}{$u_n = w_n - 0.5$} 
          %     & 10 &-0.0314  & 0.5577\\
          %     & 100 &  & 0.7150\\
          %     & 1000 & & 0.7826\\
          % \hline
          % \hline
          % $(w_n)$ comes from the Svensson Attractor \\ (for $a = -1.7; b = 1.8; c = -1.9; d = -0.4$): & & & \\
          % {$w_{n+1}= \begin{bmatrix} d\sin(aw_{n,x}) - \sin(bw_{n,y}) \\ 
          %                                                             c\cos(aw_{n,x})+\cos(bw_{n,y}) \end{bmatrix}$} & & & \\
          %   \multirow{3}{*}{$u_n = w_n-\overline{w}$}
          %       & 10 & -0.2243 & 0.49698 \\
          %       & 100 & &  0.63310 \\
          %       & 1000 & & 0.71151 \\
           \bottomrule
                \end{tabular}} %\ednote{http://paulbourke.net/fractals/peterdejong/}
 \caption{Following~\cite{manjunath2021universal}, we determine the multidimensional correlation coefficients $\rho$ to evidence a general reduction in the functional complexity of the map $G_T$ emerging due to a RNN. Rows correspond to distinct dynamical systems considered in the experiments
      The second column corresponds to the amount of artificial neurons (the dimension of $X$) in the implemented RNN.
      The last two columns comprise numerical estimates of $\rho$ for the relevant vectors. The linear association between $u_n$ vs $T(u_{n})$, as well as the linear relationship $\begin{bmatrix}
              x_{n-1}\\
              x_n
          \end{bmatrix}$ vs $G_T \left( \begin{bmatrix}
              x_{n-1}\\
              x_n
          \end{bmatrix}\right)$ may be contrasted by the reader. 
          Finally, the RNN's dimension (from~\eqref{Seq_RNN}\ednote{Is this reference correct?}) is provided to exhibiting that an increase in dimension of the driven system will 
typically result in a map $G_T$ with less functional complexity.}.
      \end{table}\label{tbl_attractorsPearson}
\end{center}
